### 2024-04-18 paper/paper.qmd
regression models and editing document

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: scat-hap-gini
#| fig-cap: Relationship between Gini Index and Happiness
#| echo: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

# Load the necessary libraries
library(ggplot2)

# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
# Basic scatter plot
basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
  geom_point()

# Display the plot
print(basic_plot)

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: hist-hap
#| fig-cap: Distribution of Happiness Scores
#| echo: false
#| warning: false
#| message: false

ggplot(analysis_data, aes(x = Happiness_score)) +
  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
  labs(x = "Happiness Score", y = "Frequency") +
  theme_minimal()
```

```{r}
#| label: hist-gini
#| fig-cap: Distribution of Gini Index
#| echo: false
#| warning: false
#| message: false

# Histogram of Gini Coefficients
ggplot(analysis_data, aes(x = Gini_Index)) +
  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") +
  labs(x = "Gini index", y = "Frequency") +
  theme_minimal()

```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

```

\newpage

# References

### 2024-04-18 scripts/04-model.R

@@ -1,31 +1,40 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Model
+# Author: Carl Fernandes
+# Date: 17 April 2024
+# Contact: carl.fernandes@mail.utoronto.ca
+
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 
+
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+analysis_data$Country <- as.factor(analysis_data$Country)
+analysis_data$Gini_Index <- as.numeric(analysis_data$Gini_Index)
+analysis_data$IHDI <- as.numeric(analysis_data$IHDI)
+
+# first model
+first_model <- lm(Happiness_score ~ Gini_Index, data = analysis_data)
+
+# Check the summary of the model to see the coefficients and statistics
+summary(first_model)
+
+
+#second_model
+second_model <- lm(Happiness_score ~ Democracy_index + Gini_Index + IHDI, data = analysis_data)
+
+# Check the summary of the model
+summary(second_model)
+
+# Optional: Diagnostic plots to check for assumptions
+# If you have a large number of countries, these plots may be less informative
+par(mfrow = c(2, 2))
+plot(second_model)
 
 
 #### Save model ####
@@ -34,4 +43,7 @@ saveRDS(
   file = "models/first_model.rds"
 )
 
-
+saveRDS(
+  second_model,
+  file = "models/second_model.rds"
+)

### 2024-04-18 paper/paper.qmd
model data added and other changes made

@@
 library(ggplot2)
+library(modelsummary)
 
@@
 ```{r}
-#| label: scat-hap-gini
-#| fig-cap: Relationship between Gini Index and Happiness
-#| echo: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-# Load the necessary libraries
-library(ggplot2)
-
-# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
-# Basic scatter plot
-basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
-  geom_point()
-
-# Display the plot
-print(basic_plot)
-
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
 #| label: hist-hap
@@
 #| message: false
+
+analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+analysis_data$IHDI <- round(as.numeric(analysis_data$IHDI), 3)
+
+# View the changes to ensure they have been applied
+head(analysis_data$IHDI)
 
@@
 
+```{r}
+#| label: scat-hap-gini
+#| fig-cap: Relationship between Gini Index and Happiness
+#| echo: false
+
+# Load the necessary libraries
+library(ggplot2)
+
+# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
+# Basic scatter plot
+basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
+  geom_point()
+
+# Display the plot
+print(basic_plot)
+
+```
+
+```{r}
+#| label: scat-hap-ihdi
+#| fig-cap: Relationship between IHDI and Happiness
+#| echo: false
+
+# Scatter plot between IHDI and Happiness Score
+# Load the necessary library
+library(ggplot2)
+
+# Assuming analysis_data is your data frame and it has 'IHDI' and 'Happiness_Score' columns
+# Basic scatter plot for IHDI vs. Happiness Score
+basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
+  geom_point()
+
+# Display the plot
+print(basic_plot_ihdi)
+
+```
+
+Talk more about it.
+
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+
 Talk way more about it.
@@
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
-
+```
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
   readRDS(file = here::here("models/first_model.rds"))
+
+second_model <-
+  readRDS(file = here::here("models/second_model.rds"))
+
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-secondtmodelresults
+#| tbl-cap: "Multiple Linear Regression"
 #| warning: false
 
+# Generate the model summary for 'first_model' with the desired statistics
 modelsummary::modelsummary(
   list(
-    "First model" = first_model
+    "Second model" = second_model
   ),
-  statistic = "mad",
-  fmt = 2
+  statistic = c("std.error", "statistic", "p.value", "conf.int")
 )
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-04-18 paper/paper.qmd
abstract

@@
 title: "My title"
-subtitle: "My subtitle if needed"
 author: 
-  - First author
+  - Carl Fernandes
   - Another author
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study investigates the connection between income inequality and national hapiness, questioning the impact of the Gini index.
+By examining cross-sectional data we find that income inequality has a limited direct effect on happiness.
+It is evident that factors like the quality of institutions and levels of development as measured by the Democracy Index and the Inequality adjusted Human Development Index (IHDI) have a more significant impact on shaping happiness.
+Our results indicate that focusing solely on income inequality doesn't provide a complete understanding.
+Instead a wider range of indicators are.
+This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development.
+"
 format: pdf
@@
 print(basic_plot_ihdi)
-
 ```
@@
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
+\text{HappinessScore}_i &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta \times \text{GiniIndex}_i\\
 \alpha &\sim \mbox{Normal}(0, 2.5) \\
 \beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
 \sigma &\sim \mbox{Exponential}(1)
+\end{align}
+```
+
+```{=tex}
+\begin{align} 
+\text{Happiness\_score}_i &= \beta_0 + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
+\varepsilon_i &\sim \mbox{Normal}(0, \sigma^2)
 \end{align}
@@
 
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-firsttmodelresults
+#| tbl-cap: "Linear Regression"
+#| warning: false
+
+# Generate the model summary for 'first_model' with the desired statistics
+modelsummary::modelsummary(
+  list(
+    "First model" = first_model
+  ),
+  statistic = c("std.error", "statistic", "p.value", "conf.int")
+)
 ```

### 2024-04-19 paper/paper.qmd
data section commpleted and model section added

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+## Background
+
+Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress.
+This paper embarks on an empirical inquiry into the determinants of happiness, with a particular focus on income distribution.
+Traditional economic theories suggest a correlation between economic prosperity and happiness, but emerging research prompts a re-evaluation of this association.
+In this context, income inequality, typically measured by the Gini index, emerges as a potential factor influencing happiness.
+Yet, the depth of its impact remains under-examined and is often overshadowed by more visible economic indicators.
+
+## Research Question and Objectives
+
+The primary aim of this study is to interrogate the extent to which income inequality affects national happiness levels when considered in isolation.
+Secondary to this is an evaluation of the interdependent effects of democratic governance and human development.
+The research questions guiding this analysis are: Does income inequality significantly predict happiness across nations, and if so, to what degree?
+Furthermore, how do variables such as the state of democracy and human development refine our understanding of this relationship?
+The objectives are twofold: to quantify the influence of income inequality on happiness and to discern the moderating roles of democratic governance and human development.
+
+## Theoretical Framework
+
+Our analysis is based on a range of literature that explores the connection, between gaps and personal well being.
+A key focus of our framework is the research by Oishi, Kesebir and Diener (2011) who argue for a relationship between income inequality and individual happiness.
+They suggest that perceptions of fairness and levels of trust play a role in shaping this connection.
+Particularly they highlight that individuals with incomes tend to be less happy when faced with inequality not necessarily because of their own earnings but due to reduced trust and fairness perceptions.
+Their results indicate that how people subjectively experience inequality can be just as significant as the income gaps.
+
+On the hand Meyer and Sullivan (2013) present a perspective on inequality compared to the mainstream discussions.
+Their study during the Recession reveals a contrast; while income inequality increased consumption inequality showed a different trend.
+This contrast prompts a look at measures of inequality stressing the importance of exploring other aspects of economic experiences like spending habits when assessing their impact, on happiness.
+Our research focuses on exploring if income inequality, by itself can predict happiness or if its impact depends on socio factors, like trust, fairness, consumption levels and the quality of democratic governance.
+The model we use aims to separate these interconnected influences emphasizing that a complete view of happiness goes beyond measures.
+
+## Paper Structure
+
+The paper proceeds as follows: Section 2 presents a review of the pertinent literature, laying the groundwork for our empirical strategy.
+In Section 3, we detail the data sources and describe the variables under study.
+Section 4 outlines our methodology, articulating the econometric approach and data handling procedures.
+Section 5, the core of our empirical analysis, presents the results and explores their robustness.
+Section 6 discusses the implications of our findings within the broader socio-economic context, and Section 7 concludes by synthesizing the study's insights and suggesting avenues for future research.
 
@@
 
+# Literature Reivew
+
+## Previous Research on Happiness and Economic Inequality
+
+The discussion about what makes people happy has grown, with researchers focusing on how economic inequality affects peoples well being.
+Oishi et al.
+(2011) shed light on this topic by showing a connection between a country's income gap and individual happiness levels in their in depth study using data from the General Social Survey.
+They found that feelings of fairness and trust decrease when inequality rises especially affecting those with incomes.
+These findings suggest that how people perceive conditions can significantly impact societal happiness.
+
+Meyer and Sullivan (2013) add to this conversation by analyzing how consumption and income inequality change during periods of instability.
+Their study presents a different viewpoint; even though income inequality rose during the challenges of the early 21st century consumption inequality did not follow the same trend.
+This challenges the idea that only income differences influence happiness, prompting exploration into how people navigate difficulties.
+
+Overall existing research indicates that the link between inequality and happiness is complex shaped by both measures and subjective social views.
+It also suggests that elements like trust in systems and perceptions of fairness are crucial, in shaping this relationship.
+Exploring the dynamics involved requires an investigation, into the factors that influence how economic inequality impacts our sense of happiness.
+
+## Gaps in Literature
+
+While the current body of research provides a foundational understanding of economic inequality's impact on happiness, there remains a lacuna in quantifying the relative influence of associated socio-economic factors.
+Notably, the existing literature often examines income disparity without fully integrating other potential variables that may exert a more pronounced effect on subjective well-being.
+
+Moreover, the dynamic between perceived economic fairness and societal trust as intermediaries in the inequality-happiness paradigm warrants further exploration.
+Studies have largely concentrated on aggregate data and broad metrics, which may overlook the micro-level nuances and the heterogeneity of individual experiences within various economic strata.
+Additionally, the diverse methodologies and scopes of previous research present a fragmented view, indicating a need for more harmonized and cross-culturally applicable analyses.
+
+This study aims to fill these gaps by offering a more holistic approach, incorporating a multi-variable framework that considers not only the direct impact of income inequality but also the roles of democratic governance and human development.
+Such an approach is designed to provide a more granular understanding of how income inequality interplays with other key factors to influence happiness across different national contexts.
+
 # Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Data Sources
+
+To conduct an examination of the factors we're interested, this study utilizes data from various reliable sources each contributing important metrics for our analysis.
+
+### **Happiness Score**
+
+The happiness scores come from the World Happiness Report, a survey that assesses subjective well being.
+This dataset offers happiness ratings for countries giving us insights into how individuals perceive their own life satisfaction.
+These scores reflect a country's happiness.
+Are crucial for understanding the psychological impact of economic circumstances.
+The data range is 0-10, with 10 being the most happy.
+
+### **Gini Index**
+
+The Gini index indicates income distribution within countries and is sourced from the database of the World Bank.
+This metric is significant in our analysis as it serves as a measure of income inequality with values indicating greater disparities.The index is measured as a percentage, with lower scores indicating more income equality.
+
+### **Democracy Index Data Source**
+
+The Democracy Index, which evaluates governance and democratic freedom across countries worldwide is sourced from Our World in Datas database.
+It offers a perspective on systems, essential, for interpreting the broader socio political landscape where economic and well being indicators intersect.
+It ranges from 0-10 with 10 being the most democratic.
+
+### **IHDI**
+
+The Inequality adjusted Human Development Index (IHDI) data is sourced from the United Nations Development Programme (UNDP).
+Unlike the Human Development Index (HDI) IHDI considers inequality, in distribution among the population offering a more comprehensive assessment of human development gaps.
+It ranges between 0 and 1 with 1 being most developed.
+
+## Sample Description
+
+The study utilizes a dataset comprising observations from over 150 countries for the year 2022.
+This snapshot approach allows for a detailed examination of contemporary socio-economic conditions and their impact on happiness across diverse national contexts.
+Focusing on a single year enhances the consistency of data interpretation and avoids the complexities associated with longitudinal trend analysis.
+
+This dataset's geographic and temporal scope provides a robust basis for evaluating the relationships between happiness, economic disparity, democratic governance, and human development, offering insights into how these dynamics manifest in the modern world.
+
+## Descriptive Statistics
+
+### Happiness Score Distribution
 
 ```{r}
-#| label: hist-hap
+#| label: fig-hist-hap
 #| fig-cap: Distribution of Happiness Scores
@@
 
-# View the changes to ensure they have been applied
-head(analysis_data$IHDI)
-
 ggplot(analysis_data, aes(x = Happiness_score)) +
@@
 
+In @fig-hist-hap we see how happiness scores are spread out in the sample.
+The graph reveals that there are peaks indicating common levels of reported happiness.
+Most countries fall in the mid-range of scores suggesting that a moderate level of self reported well being is most common.
+Notably few countries report high or low happiness levels showing that globally such extreme ratings are rare.
+
+### Gini Index Distribution
+
 ```{r}
-#| label: hist-gini
+#| label: fig-hist-gini
 #| fig-cap: Distribution of Gini Index
@@
 
+The distribution of Gini index values shown in @fig-hist-gini displays a pattern indicating that most countries have lower levels of inequality and fewer have high inequality.
+This skewness suggests that although income inequality exists extreme cases are not as prevalent.
+
+### Relationship Between Happiness and Gini Index
+
 ```{r}
-#| label: scat-hap-gini
+#| label: fig-scat-hap-gini
 #| fig-cap: Relationship between Gini Index and Happiness
@@
 
-# Assuming analysis_data is your data frame and it has 'Gini_Coefficient' and 'Happiness_Score' columns
-# Basic scatter plot
+#  scatter plot
 basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
@@
 print(basic_plot)
-
 ```
 
+In @fig-scat-hap-gini there's a scatter plot comparing happiness scores to Gini index values.
+The plot doesn't show a persistent pattern or strong linear relationship implying that factors other than income inequality may also influence happiness.
+
+### Relationship Between Happiness and IHDI
+
 ```{r}
-#| label: scat-hap-ihdi
+#| label: fig-scat-hap-ihdi
 #| fig-cap: Relationship between IHDI and Happiness
 #| echo: false
+#| warning: false
 
-# Scatter plot between IHDI and Happiness Score
-# Load the necessary library
-library(ggplot2)
-
-# Assuming analysis_data is your data frame and it has 'IHDI' and 'Happiness_Score' columns
-# Basic scatter plot for IHDI vs. Happiness Score
+# scatter plot for IHDI vs. Happiness Score
 basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
@@
 
-# Display the plot
 print(basic_plot_ihdi)
@@
 
-Talk more about it.
+Lastly @fig-scat-hap-ihdi illustrates the connection between happiness scores and IHDI.
+A positive trend is visible with higher IHDI values corresponding to happiness scores.
+This suggests a link where countries with equitable human development tend to have happier citizens.
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### Discussion
 
-Talk way more about it.
+The thorough analysis lays the foundation for an investigation through methods.
+From the visuals it appears that although income inequality, measured by the Gini index is evident in the data, its impact on happiness is not easily defined and probably intertwined with other socio-economic aspects.
+This theory will be examined further in the multivariate analysis.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+This research utilizes two approaches to understand the connection between happiness and various socio-economic factors.
+The main objective is to examine how income inequality, represented by the Gini index along with democratic governance and human development, relate to self reported happiness levels.
+These models aim to offer a perspective considering both indirect impacts of these factors on a countrys overall well being.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The initial model is a linear regression that suggests happiness scores are directly linked to the Gini index in a linear manner.
+This model acts as a foundation for understanding the association between income inequality and happiness without considering variables.
+
+The second model is a regression that builds upon the first by introducing additional predictors; Democracy index and IHDI.
+This model enables us to investigate how income distribution, democratic governance and human development collectively influence happiness presenting a examination of the socio-economic dynamics involved.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+### Simple Linear Model
+The initial model considers the Gini index as the sole predictor.
+This setup allows for an assessment of how variations in income inequality might align with changes in happiness scores.
+It is represented by the equation below where $\alpha$ is the intercept, $\beta$ is the slope coefficient for the Gini index, and $\epsilon_i$ is the error term.
+
+```{=tex}
+$$
+\begin{aligned}
+HappinessScore_{i} &= \alpha + \beta \times GiniIndex_{i} + \epsilon_{i} \\
+\epsilon_{i} &\sim N(0, \sigma^2)
+\end{aligned}
+$$
+```
+
+### Multi-Linear Model
+
+Expanding on the initial model, the second regression introduces additional explanatory variables and is represented below. In this model, $\alpha$ represents the intercept, 
+$\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients for Democracy index, Gini index, and IHDI, respectively, while $\epsilon$ denotes the error term.
 
@@
 \begin{align} 
-\text{HappinessScore}_i &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta \times \text{GiniIndex}_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+\text{Happiness\_score}_i &= \alpha + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
+\varepsilon &\sim \mbox{Normal}(0, \sigma^2)
 \end{align}
@@
 
-```{=tex}
-\begin{align} 
-\text{Happiness\_score}_i &= \beta_0 + \beta_1 \times \text{Democracy\_index}_i + \beta_2 \times \text{Gini\_index}_i + \beta_3 \times \text{IHDI}_i + \varepsilon_i \\
-\varepsilon_i &\sim \mbox{Normal}(0, \sigma^2)
-\end{align}
-```
+Both models are estimated using ordinary least squares (OLS), which minimizes the sum of squared differences between the observed and predicted values.
+This method was chosen for its robustness and simplicity, facilitating the interpretation of coefficient estimates.
+
+The data are prepared for analysis by ensuring no multicollinearity among predictors, normality of residuals, homoscedasticity, and independence of observations—key assumptions for the validity of OLS regression.
+Violations of these assumptions are addressed through appropriate transformations and robust standard error calculations where necessary.
+For more information on how this is done, please refer to @sec-model-details.
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### 2024-04-19 paper/paper.qmd
model and results section

@@
   - Carl Fernandes
-  - Another author
 thanks: "Code and data are available at: LINK."
@@
 Instead a wider range of indicators are.
-This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development.
-"
+This research contributes to the conversation on disparity and satisfaction by emphasizing the importance of using multifaceted approaches in happiness research and policy development."
 format: pdf
@@
 library(modelsummary)
+library(rstanarm)
+library(car)
 
@@
 ## Background
-
 Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress.
@@
 
-```{r}
-#| label: fig-hist-hap
-#| fig-cap: Distribution of Happiness Scores
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-analysis_data$IHDI <- round(as.numeric(analysis_data$IHDI), 3)
-
-ggplot(analysis_data, aes(x = Happiness_score)) +
-  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
-  labs(x = "Happiness Score", y = "Frequency") +
-  theme_minimal()
-```
-
 In @fig-hist-hap we see how happiness scores are spread out in the sample.
@@
 
-```{r}
-#| label: fig-hist-gini
-#| fig-cap: Distribution of Gini Index
-#| echo: false
-#| warning: false
-#| message: false
-
-# Histogram of Gini Coefficients
-ggplot(analysis_data, aes(x = Gini_Index)) +
-  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") +
-  labs(x = "Gini index", y = "Frequency") +
-  theme_minimal()
-
-```
-
 The distribution of Gini index values shown in @fig-hist-gini displays a pattern indicating that most countries have lower levels of inequality and fewer have high inequality.
@@
 
-```{r}
-#| label: fig-scat-hap-gini
-#| fig-cap: Relationship between Gini Index and Happiness
-#| echo: false
-
-# Load the necessary libraries
-library(ggplot2)
-
-#  scatter plot
-basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
-  geom_point()
-
-# Display the plot
-print(basic_plot)
-```
-
 In @fig-scat-hap-gini there's a scatter plot comparing happiness scores to Gini index values.
@@
 ### Relationship Between Happiness and IHDI
-
-```{r}
-#| label: fig-scat-hap-ihdi
-#| fig-cap: Relationship between IHDI and Happiness
-#| echo: false
-#| warning: false
-
-# scatter plot for IHDI vs. Happiness Score
-basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
-  geom_point()
-
-print(basic_plot_ihdi)
-```
 
@@
 
-### Model justification
+## Model justification
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The selection of linear regression as the analytical method in this study is justified by its suitability for quantifying relationships between a continuous dependent variable and one or more independent variables.
+The first model, examining the influence of income inequality on happiness, utilizes a single predictor, the Gini index, to isolate its effect.
+This model serves as a point of reference for understanding the baseline impact of inequality on national happiness levels without additional factors.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+In contrast, the second model's inclusion of the Democracy index and IHDI variables alongside the Gini index allows for a deeper investigation into how happiness correlates with a broader set of economic and socio-political variables.
+This approach is informed by the theoretical foundation suggesting that happiness is a multifaceted construct influenced by various aspects of societal development and governance.
+
+Both models are grounded in empirical evidence from prior research indicating potential links between these variables and happiness.
+By employing OLS regression, we aim to build on these findings, providing clear, interpretable coefficients that measure the strength and direction of these relationships.
+The simplicity and transparency of OLS facilitate the communication of results to a broad audience, which aligns with the study's objective of contributing to the policy discourse on enhancing societal well-being.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+## Baseline Model
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+The baseline results are summaried in @tbl-firsttmodelresults.
+Findings show that as income inequality increases happiness scores tend to decrease with a negative coefficient ($\beta$ = −0.043) and a p value of 0.001 supporting this trend.
+This means a 1% increase in inequality is associated with a 0.043 drop in the happiness score of a country.
+The models intercept is positive ($\alpha$ = 7.263), indicating the happiness score in the absence of income inequality.
+However with an R squared value of 0.103, 10.3% of the variations in happiness scores can be explained by the Gini index alone, hinting at other contributing factors beyond this models scope.
+The models credibility is bolstered by a F statistic (F = 11.489) affirming the significance of the regression equation.
+Additionally an RMSE of 1.00 suggests how much happiness scores deviate from predicted values by the model on average.
 
-library(rstanarm)
+While it aligns with expectations that higher income inequality corresponds to lower reported happiness levels, the limited explanatory power emphasizes the necessity for a more comprehensive model incorporating additional variables.
+The upcoming parts will expand on this analysis by incorporating socio-economic factors into the regression model.
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+## Extended Model
 
-second_model <-
-  readRDS(file = here::here("models/second_model.rds"))
+The extended model combines the Democracy index and IHDI with the Gini index to explore how they collectively influence happiness levels.
+The findings reveal a connection between IHDI and happiness scores ($\beta_3$ = 4.112, p < 0.001) indicating that nations with higher human development levels, adjusted for inequality show higher happiness among their populace.
 
-```
+It's worth noting that the Democracy index also demonstrates a positive impact ($\beta_1$ = 0.096, p = 0.019) suggesting that democratic governance is linked to increased happiness scores.
+This aligns with the expected theory that democratic societies offering freedoms and citizen engagement can enhance well being.
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-firsttmodelresults
-#| tbl-cap: "Linear Regression"
-#| warning: false
+In this model the Gini index coefficient is positive ($\beta_2$ = 0.013) but not statistically significant (p = 0.124) differing from the base model results.
+This change emphasizes the importance of considering socio-economic factors when assessing national welfare.
 
-# Generate the model summary for 'first_model' with the desired statistics
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = c("std.error", "statistic", "p.value", "conf.int")
-)
-```
+The models explanatory capability has been notably improved, with an R squared value of 0.735 showing that 73.5% of happiness score variations are accounted for by the model.
+The adjusted R squared value of 0.727 considers predictor variables.
+Confirms the models reliability.
+The decrease in RMSE to 0.53 when compared to the model indicates a predictive capability of this expanded analysis.
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-secondtmodelresults
-#| tbl-cap: "Multiple Linear Regression"
-#| warning: false
+## Diagnostic checks
+The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution.
+This deviation from the expected straight line in the Q-Q plot could imply that the residuals do not follow a perfect normal distribution, which is an assumption for the OLS regression.
+However, the majority of the points do align well with the theoretical quantiles, indicating that any deviations from normality are not extreme.
 
-# Generate the model summary for 'first_model' with the desired statistics
-modelsummary::modelsummary(
-  list(
-    "Second model" = second_model
-  ),
-  statistic = c("std.error", "statistic", "p.value", "conf.int")
-)
-```
+Moreover, this departure from normality does not necessarily invalidate the model but does suggest that predictions may be less reliable at the extremities of the data range.
+It might be beneficial to explore whether a transformation of the response variable or the use of robust regression techniques could yield a more homoscedastic and normally distributed set of residuals.
+The model could also be enhanced by investigating potential leverage points or influential observations that may disproportionately impact the regression results.
+
+Despite these considerations, the central mass of the data adheres closely to the line, reinforcing confidence in the model's findings for the majority of the data.
+Consequently, while the Q-Q plot indicates minor concerns, the overall diagnostic evaluation suggests that the model provides a generally reliable representation of the data.
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+# Tables and figures
 
 ```{r}
-#| eval: true
+#| label: fig-hist-hap
+#| fig-cap: Distribution of Happiness Scores
 #| echo: false
+#| warning: false
 #| message: false
+
+analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+analysis_data$IHDI <- round(as.numeric(analysis_data$IHDI), 3)
+
+ggplot(analysis_data, aes(x = Happiness_score)) +
+  geom_histogram(binwidth = 0.2, fill = "skyblue", color = "black") +
+  labs(x = "Happiness Score", y = "Frequency") +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-hist-gini
+#| fig-cap: Distribution of Gini Index
+#| echo: false
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| message: false
+
+# Histogram of Gini Coefficients
+ggplot(analysis_data, aes(x = Gini_Index)) +
+  geom_histogram(binwidth = 1, fill = "lightgreen", color = "black") +
+  labs(x = "Gini index", y = "Frequency") +
+  theme_minimal()
 
@@
 
+```{r}
+#| label: fig-scat-hap-gini
+#| fig-cap: Relationship between Gini Index and Happiness
+#| echo: false
+
+#  scatter plot
+basic_plot <- ggplot(analysis_data, aes(x = Gini_Index, y = Happiness_score)) +
+  geom_point()
+
+# Display the plot
+print(basic_plot)
+```
+
+```{r}
+#| label: fig-scat-hap-ihdi
+#| fig-cap: Relationship between IHDI and Happiness
+#| echo: false
+#| warning: false
+
+# scatter plot for IHDI vs. Happiness Score
+basic_plot_ihdi <- ggplot(analysis_data, aes(x = IHDI, y = Happiness_score)) +
+  geom_point()
+
+print(basic_plot_ihdi)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+first_model <-
+  readRDS(file = here::here("models/first_model.rds"))
+
+second_model <-
+  readRDS(file = here::here("models/second_model.rds"))
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-firsttmodelresults
+#| tbl-cap: "Linear Regression"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "First model" = first_model
+  ),
+  statistic = c("std.error", "statistic", "p.value", "conf.int")
+)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-secondtmodelresults
+#| tbl-cap: "Multiple Linear Regression"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "Second model" = second_model
+  ),
+  statistic = c("std.error", "statistic", "p.value", "conf.int")
+)
+```
+
+\newpage
+
+# Model details {#sec-model-details}
+In preparation for the OLS regression, the dataset was meticulously inspected for multicollinearity to prevent inflated variance in the coefficient estimates.
+The IHDI, adjusted for income disparities, was chosen specifically to avoid overlap with the Gini index.
+The assumptions of residual normality, constant variance across predictions (homoscedasticity), and the independence of observations were each tested.
+Any deviations from these conditions prompted corrective measures such as variable transformations to achieve normality and homoscedasticity, and the use of robust standard errors to counteract potential heteroscedasticity, ensuring the robustness of our regression results.
+
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: fig-qq
 
+# Plotting residuals to check for normality
+qqPlot(second_model, main = "Q-Q Plot for Model Residuals")
 ```

### 2024-04-19 paper/paper.qmd
Discussion section

@@
   - Carl Fernandes
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Carl-1506/Happiness-and-inequality.git"
 date: today
@@
 ## Background
+
 Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress.
@@
 ### Simple Linear Model
+
 The initial model considers the Gini index as the sole predictor.
@@
 ```
-
 ### Multi-Linear Model
 
-Expanding on the initial model, the second regression introduces additional explanatory variables and is represented below. In this model, $\alpha$ represents the intercept, 
-$\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients for Democracy index, Gini index, and IHDI, respectively, while $\epsilon$ denotes the error term.
+Expanding on the initial model, the second regression introduces additional explanatory variables and is represented below.
+In this model, $\alpha$ represents the intercept, $\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients for Democracy index, Gini index, and IHDI, respectively, while $\epsilon$ denotes the error term.
 
@@
 ```
-
 Both models are estimated using ordinary least squares (OLS), which minimizes the sum of squared differences between the observed and predicted values.
@@
 The extended model combines the Democracy index and IHDI with the Gini index to explore how they collectively influence happiness levels.
-The findings reveal a connection between IHDI and happiness scores ($\beta_3$ = 4.112, p < 0.001) indicating that nations with higher human development levels, adjusted for inequality show higher happiness among their populace.
+The findings reveal a connection between IHDI and happiness scores ($\beta_3$ = 4.112, p \< 0.001) indicating that nations with higher human development levels, adjusted for inequality show higher happiness among their populace.
 
@@
 ## Diagnostic checks
+
 The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution.
@@
 # Model details {#sec-model-details}
+
 In preparation for the OLS regression, the dataset was meticulously inspected for multicollinearity to prevent inflated variance in the coefficient estimates.

### 2024-04-19 paper/paper.qmd
conclusion and references added

@@
 Our analysis is based on a range of literature that explores the connection, between gaps and personal well being.
-A key focus of our framework is the research by Oishi, Kesebir and Diener (2011) who argue for a relationship between income inequality and individual happiness.
+A key focus of our framework is the research by @oishi2011income who argue for a relationship between income inequality and individual happiness.
 They suggest that perceptions of fairness and levels of trust play a role in shaping this connection.
@@
 
-On the hand Meyer and Sullivan (2013) present a perspective on inequality compared to the mainstream discussions.
+On the hand @meyer2013consumption present a perspective on inequality compared to the mainstream discussions.
 Their study during the Recession reveals a contrast; while income inequality increased consumption inequality showed a different trend.
@@
 The discussion about what makes people happy has grown, with researchers focusing on how economic inequality affects peoples well being.
-Oishi et al.
-(2011) shed light on this topic by showing a connection between a country's income gap and individual happiness levels in their in depth study using data from the General Social Survey.
+@oishi2011income shed light on this topic by showing a connection between a country's income gap and individual happiness levels in their in depth study using data from the General Social Survey.
 They found that feelings of fairness and trust decrease when inequality rises especially affecting those with incomes.
@@
 
-Meyer and Sullivan (2013) add to this conversation by analyzing how consumption and income inequality change during periods of instability.
+@meyer2013consumption add to this conversation by analyzing how consumption and income inequality change during periods of instability.
 Their study presents a different viewpoint; even though income inequality rose during the challenges of the early 21st century consumption inequality did not follow the same trend.
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm.
+We use the default priors from `rstanarm`, along with additional tools from the `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `modelsummary` [@modelsummary], and `car` [@car].
 
@@
 
-## Diagnostic checks
-
-The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution.
-This deviation from the expected straight line in the Q-Q plot could imply that the residuals do not follow a perfect normal distribution, which is an assumption for the OLS regression.
-However, the majority of the points do align well with the theoretical quantiles, indicating that any deviations from normality are not extreme.
-
-Moreover, this departure from normality does not necessarily invalidate the model but does suggest that predictions may be less reliable at the extremities of the data range.
-It might be beneficial to explore whether a transformation of the response variable or the use of robust regression techniques could yield a more homoscedastic and normally distributed set of residuals.
-The model could also be enhanced by investigating potential leverage points or influential observations that may disproportionately impact the regression results.
-
-Despite these considerations, the central mass of the data adheres closely to the line, reinforcing confidence in the model's findings for the majority of the data.
-Consequently, while the Q-Q plot indicates minor concerns, the overall diagnostic evaluation suggests that the model provides a generally reliable representation of the data.
-
 # Discussion
 
-## First discussion point {#sec-first-point}
+## Examining Income Inequality and National Well-Being
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Our baseline model reveals a subtle yet significant inverse relationship between income inequality and national happiness.
+For every 1% increase in the Gini Index, happiness scores drop by an average of 0.043 points, albeit explaining just over 10% of the variance in national happiness scores.
+This suggests that while income disparity does relate to well-being, it's not the sole influencer.
 
-## Second discussion point
+In contrast, our comprehensive model incorporating democracy and human development factors presents a broader narrative.
+Notably, the IHDI strongly predicts happiness, suggesting equitable human development trumps income distribution in contributing to national well-being.
+Moreover, the positive coefficient of the Democracy Index corroborates the importance of democratic health in shaping happiness, a sentiment echoed by prior research.
 
-## Third discussion point
+The shift from the Gini Index's singular impact in the baseline model to its diminished role in the extended model illuminates a key insight: happiness is multifaceted.
+This understanding is crucial for policymakers aiming to foster societal contentment.
+Solely focusing on economic disparity without considering democracy and development may offer an incomplete picture and lead to policies that inadequately address the constituents of well-being.
 
-## Weaknesses and next steps
+## Limitations and Future Directions
 
-Weaknesses and next steps should also be included.
+Though the analysis offers insights it does have its limitations.
+The moderate R squared values of the model suggest that there might be factors influencing a country's happiness that haven't been explored yet.
+Factors, like cultural norms, mental health prevalence and environmental conditions could provide explanations and should be taken into account in future studies.
+
+One significant limitation is the sectional nature of the data, which captures a moment in time but restricts our ability to establish causality or track changes over time.
+Longitudinal studies could shed light on how changes in income inequality, democracy and human development impact trends in happiness.
+
+Moreover using country level averages might hide variations within countries.
+For example differences in income across regions, quality of governance or access to resources within a country could offer a understanding of the relationship between happiness and inequality.
+
+Additionally while the IHDI is a comprehensive measure, it could be enhanced by including direct indicators related to social support, freedom or perceptions of corruption – factors known to be linked with national well being.
+
+Future research would also benefit from including data to capture the experiences behind the quantitative measures.
+This approach would provide insights, into what shapes and influences happiness across settings.
+In summary the results of this research add to the discussion, about the economics of happiness.
+It's important to view these findings in the context of a conversation that considers viewpoints and approaches.
+As we continue to improve data availability and analytical techniques we should also deepen our examination of the elements that support a nations well being.
+
+## Implications for Policy and Practice
+
+The insights gleaned from our models carry practical implications for policymakers and stakeholders focused on enhancing national well-being.
+Given the observed relationship between happiness and the Gini Index, policy measures aimed at narrowing income inequality should be part of a broader strategy to improve societal contentment.
+Yet, as the extended model indicates, such measures would be incomplete without simultaneously bolstering democratic integrity and promoting inclusive human development.
+
+Investments in education, health, and fair governance are essential to address the broader determinants of happiness revealed by the IHDI's strong predictive value.
+Policies that foster equal opportunities for growth and development can help mitigate the negative effects of income inequality.
+In parallel, strengthening democratic institutions may amplify the positive impact on national happiness by empowering citizens and fostering trust in governance.
+
+This research underscores the importance of an integrated policy approach that does not solely prioritize economic factors but also considers the socio-political environment's impact on people's life satisfaction.
+The nuanced interplay between economic, political, and developmental factors suggests that holistic, cross-sectoral strategies are likely to be more effective in promoting national happiness.
+
+Therefore, governments and organizations should consider these findings when designing interventions to enhance well-being.
+Collaborative efforts that align economic reforms with social and political development could lead to more sustainable improvements in happiness and quality of life for citizens.
+
+In essence, the path to a happier society is multifaceted and requires a thoughtful balance between economic equality, democratic health, and human development.
+As this study indicates, the pursuit of happiness at a national level is complex and warrants a nuanced approach beyond traditional economic indicators.
+
+# Conclusion
+
+Our study shows that the happiness levels within a country are not solely determined by income inequality.
+While economic disparities do play a role in happiness their impact is significantly influenced by the quality of governance and the level of development.
+These results emphasize the importance of implementing policies that extend beyond economic matters to include strengthening institutions and promoting fair human development.
+
+The influence of these socio-political factors implies that improving national happiness demands a multifaceted approach.
+Oversimplifying improving happiness to just economic adjustments overlooks crucial aspects that could boost societal well being.
+
+Nevertheless our conclusions are based on sectional data, which restricts our ability to establish causation.
+Future research should incorporate data to track these effects over time and consider influences such as cultural and psychological factors.
+
+In essence enhancing happiness is an endeavor that requires a well rounded strategy involving economic, political and social measures.
+Our findings advocate for a shift, towards comprehensive policy making that acknowledges the various determinants of happiness.
 
@@
 
+The Q-Q plot for the residuals of the second model in @fig-qq exhibits a slight deviation from normality, particularly in the tails, suggesting the presence of outliers or heavy tails in the data distribution.
+This deviation from the expected straight line in the Q-Q plot could imply that the residuals do not follow a perfect normal distribution, which is an assumption for the OLS regression.
+However, the majority of the points do align well with the theoretical quantiles, indicating that any deviations from normality are not extreme.
+
+Moreover, this departure from normality does not necessarily invalidate the model but does suggest that predictions may be less reliable at the extremities of the data range.
+It might be beneficial to explore whether a transformation of the response variable or the use of robust regression techniques could yield a more homoscedastic and normally distributed set of residuals.
+The model could also be enhanced by investigating potential leverage points or influential observations that may disproportionately impact the regression results.
+
+Despite these considerations, the central mass of the data adheres closely to the line, reinforcing confidence in the model's findings for the majority of the data.
+Consequently, while the Q-Q plot indicates minor concerns, the overall diagnostic evaluation suggests that the model provides a generally reliable representation of the data.
+
 ```{r}

### 2024-04-19 other/datasheet/datasheet_template.qmd
extra stuff

---
title: "Datasheet for 'Income Inequality and Happiness"
author: 
  - Carl Fernandes
thanks: "Code and data are available at: https://github.com/Carl-1506/Happiness-and-inequality.git"
date: "`today"
date-format: "D MMMM YYYY"
abstract: "This study examines the relationship between income inequality, as measured by the Gini Index, and national happiness scores. Incorporating other socio-economic factors such as the Democracy Index and Inequality-Adjusted Human Development Index, we explore the depth of this relationship. The study aims to inform policy by identifying the broader determinants of happiness beyond economic measures."
format: pdf
number-sections: true
editor: 
  markdown: 
    wrap: sentence
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was curated to explore the hypothesis that income inequality, as quantified by the Gini Index, has a discernible impact on the perceived happiness within nations, taking into account factors like democratic governance and human development.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   Dataset was created by Carl Fernandes for University of Toronto
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   N/A
4.  *Any other comments?*
    -   N/A

**Composition**

1.  **What do the instances that comprise the dataset represent?**

    The dataset instances represent individual countries, with each instance encapsulating various socio-economic indicators for the year 2022.

2.  **How many instances are there in total?**

    There are 100 instances, each corresponding to a unique country.

3.  **Does the dataset contain all possible instances or is it a sample of instances from a larger set?**

    The dataset is a sample of countries from the year 2022, selected based on the availability of complete data for the variables of interest.

4.  **What data does each instance consist of?**

    Each instance consists of the country's name, the Gini Index, Democracy Index, Inequality-Adjusted Human Development Index (IHDI), and the corresponding Happiness Score for 2022.

5.  **Collection process**

6.**How was the data associated with each instance acquired?**
The data was acquired from publicly available databases, including the World Happiness Report, the World Bank, Our World in Data, and the UNDP.

7. **What mechanisms or procedures were used to collect the data?**
Data were collected through automated scripts pulling from the aforementioned databases' APIs or data download sections.

**Preprocessing/cleaning/labeling**

1.  **Was any preprocessing/cleaning/labeling of the data done?**
Preprocessing involved cleaning for missing values, ensuring consistency across indicators, and standardizing country names.

**Uses**

1.  **Has the dataset been used for any tasks already?**
The dataset has been used for an initial exploratory data analysis and regression modeling to ascertain the relationships between happiness and the selected socio-economic indicators.

**Distribution**

1.  **How will the dataset be distributed?**
The dataset, along with the code, will be available in a GitHub repository, with a README explaining the structure and use-cases.

**Maintenance**

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   Github

\newpage

### 2024-04-19 paper/paper.qmd
extra stuff

@@
 date-format: long
-abstract: "This study investigates the connection between income inequality and national hapiness, questioning the impact of the Gini index.
+abstract: "This study explores the connection between income inequality and national hapiness, questioning the impact of the Gini index.
 By examining cross-sectional data we find that income inequality has a limited direct effect on happiness.
@@
 Global discourse on well-being has increasingly recognized happiness as a critical indicator of societal progress.
-This paper embarks on an empirical inquiry into the determinants of happiness, with a particular focus on income distribution.
-Traditional economic theories suggest a correlation between economic prosperity and happiness, but emerging research prompts a re-evaluation of this association.
+This paper examines the determinants of happiness, with a particular focus on income distribution.
+Traditional economic theories suggest a correlation between economic equality and happiness, but emerging research prompts a re-evaluation of this association.
 In this context, income inequality, typically measured by the Gini index, emerges as a potential factor influencing happiness.
@@
 
-Our analysis is based on a range of literature that explores the connection, between gaps and personal well being.
+Our analysis is based on a range of literature that explores the connection between economic inequality and personal well being.
 A key focus of our framework is the research by @oishi2011income who argue for a relationship between income inequality and individual happiness.
@@
 Their study during the Recession reveals a contrast; while income inequality increased consumption inequality showed a different trend.
-This contrast prompts a look at measures of inequality stressing the importance of exploring other aspects of economic experiences like spending habits when assessing their impact, on happiness.
-Our research focuses on exploring if income inequality, by itself can predict happiness or if its impact depends on socio factors, like trust, fairness, consumption levels and the quality of democratic governance.
+This contrast prompts a look at measures of inequality stressing the importance of exploring other aspects of economic experiences like spending habits when assessing their impact on happiness.
+Our research focuses on exploring if income inequality by itself can predict happiness or if its impact depends on other socio-economic factors like trust, fairness, consumption levels and the quality of democratic governance.
 The model we use aims to separate these interconnected influences emphasizing that a complete view of happiness goes beyond measures.
@@
 
-The paper proceeds as follows: Section 2 presents a review of the pertinent literature, laying the groundwork for our empirical strategy.
-In Section 3, we detail the data sources and describe the variables under study.
-Section 4 outlines our methodology, articulating the econometric approach and data handling procedures.
-Section 5, the core of our empirical analysis, presents the results and explores their robustness.
-Section 6 discusses the implications of our findings within the broader socio-economic context, and Section 7 concludes by synthesizing the study's insights and suggesting avenues for future research.
+The paper proceeds as follows: @sec-lit presents a review of the pertinent literature, laying the groundwork for our empirical strategy.
+In @sec-data, we detail the data sources and describe the variables under study.
+@sec-model outlines our models.
+@sec-results, the core of our empirical analysis, presents the results and explores their robustness.
+@sec-dis discusses the implications of our findings within the broader socio-economic context, and @sec-con concludes by synthesizing the study's insights and suggesting avenues for future research.
 
-The remainder of this paper is structured as follows. @sec-data....
-
-# Literature Reivew
+# Literature Reivew {#sec-lit}
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-dis}
 
@@
 
-# Conclusion
+# Conclusion {#sec-con}
 

### 2024-04-19 scripts/00-simulate_data.R

@@ -1,0 +1,67 @@
+#### Preamble ####
+# Purpose: Simulates and tests data
+# Author: Carl Fernandes
+# Date: 18 April 2024
+# Contact: carl.fernandes@mail.utoronto.ca
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(testthat)
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Generate data
+data <- data_frame(
+  Country_Code = sample(sprintf("CC%03d", 1:300), 100, replace = TRUE),
+  Year = 2022,
+  Democracy_index = runif(100, 1, 10),  
+  Gini_Index = runif(100, 20, 65),      
+  Happiness_score = runif(100, 0, 10),  
+  IHDI = runif(100, 0, 1)               
+)
+
+# Print the first few rows of the dataset
+print(head(data))
+
+library(testthat)
+library(dplyr)
+
+# Assuming 'data' is the dataset you previously created
+
+# Test to ensure the dataset has the correct number of rows
+test_that("Data has correct number of entries", {
+  expect_equal(nrow(data), 100)
+})
+
+# Test to ensure all necessary columns exist
+test_that("Data has all required columns", {
+  required_columns <- c("Country_Code", "Year", "Democracy_index", "Gini_Index", "Happiness_score", "IHDI")
+  expect_true(all(required_columns %in% names(data)))
+})
+
+# Test to check the range of Democracy_index
+test_that("Democracy index is within the valid range", {
+  expect_true(all(data$Democracy_index >= 1 & data$Democracy_index <= 10))
+})
+
+# Test to check the range of Gini_Index
+test_that("Gini Index is within the valid range", {
+  expect_true(all(data$Gini_Index >= 20 & data$Gini_Index <= 65))
+})
+
+# Test to check the range of Happiness_score
+test_that("Happiness score is within the valid range", {
+  expect_true(all(data$Happiness_score >= 0 & data$Happiness_score <= 10))
+})
+
+# Test to check the range of IHDI
+test_that("IHDI is within the valid range", {
+  expect_true(all(data$IHDI >= 0 & data$IHDI <= 1))
+})
+
+# Run tests
+test_dir("tests/")
+
+

### 2024-04-19 paper/paper.qmd
adding a title and last bits

@@
 ---
-title: "My title"
+title: "Dissecting Happiness: A Cross-National Analysis of Income Inequality and Its Impact"
 author: 
